Population structure and phenotypic variation of *Sclerotinia sclerotiorum* from dry bean in the United States

### 2017-08-21 doc/manuscript/manuscript.Rmd
Update introduction to include info about popgen

Currently, there are no known resistant cultivars of dry bean commercially available [@otto-hanson2011variation].
Organized breeding efforts have been using a common-garden approach with white mold screening nurseries in dry bean production areas across the United States with additional sites in Mexico, France, and Australia [@steadman2003identification; @steadman2004identification; @steadman2005identification; @steadman2006identification; @otto2007identification]. 
These white mold screening nurseries use no chemical or cultural treatments against *S. sclerotiorum* and employ standardized protocols for screening new cultivars for resistance to white mold. 
It has previously been shown that aggressiveness (the severity of disease symptoms on the host) is significantly different across white mold screening nursery sites in separate geographic regions [@otto-hanson2011variation].
The genetic structure and mode of reproduction in these populations, however, is currently unknown.
Understanding genetic relationships and reproduction behavior of *S. sclerotiorum* populations is beneficial for breeders seeking to develop new resistant cultivars for worldwide deployment [@mcdonald2002pathogen; @milgroom1996recombination].

Most populations of *S. sclerotiorum* are predominantly clonal and have a large degree of population fragmentation [@attanayake2012genetic; @carbone2001multilocus; @cubeta1997clonality; @ekins2011population; @kohli1998random; @kohli1995clonal]. 
Some populations are sexual [@atallah2004high; @attanayake2013sclerotinia; @sexton2004microsatellite]....

@@

<!-- ZNK: TMI -->
<!-- Sclerotia are capable of surviving more than five years in soil [@steadman1983white] and germinate carpogenically or myceliogenically to infect the plant when environmental conditions are favorable.
A carpogenically germinated sclerotia gives rise to apothecia in which ascospores are formed.
Ascospores infect senescent plant tissues that, when supplemented with cool and humid micro-climate within the plant canopy, leads to mycelial infection of the host plant.
Consequently, flowering, which coincide with canopy closure is a critical factor of disease initiation since sensing flower parts fallen on to leaves, petioles or stems provide the required energy source for plant infection.
Myceliogenically germinated sclerotia result in formation of hyphae, which also have potential for directly plant infection, though this is considered rare and less important within the disease cycle of the pathogen [@abawi1975source].
Diseased bean fields initially show scattered wilted leaves.
Symptoms on individual plants are water-soaked lesions on leaves, pods and stems, which expand over time and show signs of white fungal mycelia.
Lesions on stems and branches become necrotic as they age and dead tissues appear bleached.
Sclerotia are often produced on dead or dying plants once conditions become unfavorable for fungal growth. -->

<!-- Previous studies of *S. sclerotiorum* population biology in North America and Australia have reported clonal propagation of the pathogen [@cubeta1997clonality; @ekins2011population; @kohli1998random; @kohli1995clonal] with evidence of outcrossing observed in some areas [@atallah2004high; @kohli1998random; @malvarez2007new; @sexton2004microsatellite]. -->
<!-- One such study examined *S. sclerotiorum* population structure using microsatellite markers from potato fields in the Columbia Basin of Washington state. -->
<!-- Due to high crop diversity and high presence of the pathogen, it was important to understand the potential for outcrossing in order to develop disease management strategies. -->
<!-- Molecular analysis revealed 92% of the isolate variability was found within subpopulations [@atallah2004high]. -->
<!-- However, there was evidence of linkage equilibrium of microsatellite loci in 2 locations and illustrated the possibility of outcrossing in this region.   -->

<!-- ZNK: This is more discussion-material -->
<!-- A population study from Australian canola fields showed evidence of sexual outcrossing of *S. sclerotiorum* isolates [@sexton2004microsatellite].
Another study characterized three populations of *S. sclerotiorum* from lettuce in California (CA) and Ontario (ON), and pea/lentil in Washington (WA) [@malvarez2007new].
While all three populations were  clonal, the index of association ($I_A$) based on DNA fingerprint was closer to zero for CA than it was for the other populations.
Populations from Australian sunflower with evidence of clonal reproduction were reported, and despite populations sampled from fields separated 50 to 400 km and from both head and basal rots, all showed similar gene diversity, with authors concluding isolates representing different geographies belong to one large genetic population of *S. sclerotiorum* [@ekins2011population].  -->

<!-- SEE: Avoid sentences that start with the authors/reference. -->
<!-- @attanayake2013sclerotinia studied population differentiation of *S. sclerotiorum* isolates infecting canola from China and the USA. -->
<!-- There were no shared microsatellite multilocus genotypes (MLG) between the two populations. -->
<!-- Sexual outcrossing was detected in both populations with higher recombination rate in Chinese population than the USA population. -->
<!-- Additionally, phenotypic traits such as production of sclerotia, mycelial compatibility group (MCG), growth rate were also significantly different between the two populations. -->
<!-- Genetic diversity of *S. sclerotiorum* from soy bean, canola, sunflower, and dry bean in the North Central region of the USA was studied by [@aldrich-wolfe2015genetic]. -->
<!-- Microsatellite analysis revealed little host- specificity or genetic-isolation of *S. sclerotiorum* isolates among host crops as 93% of the observed genetic variation was explained by within-host differences. -->
<!-- Isolates from all host crops were primarily reproducing clonally while there were also evidence of genetic recombination of isolates from canola and sunflower. -->

<!-- Most of the previous studies on *S. sclerotiorum* had been conducted in the USA, Canada and Australia but not in Europe. 
In those studies, different dry bean cultivars had been rarely used as  hosts to investigate any host-pathogen relationships or specificity.
No white mold resistant dry bean cultivars exist.
Canopy architecture and growth pattern of some cultivars enable improved disease avoidance [@kolkman2002agronomic]. -->
<!-- To date, no commercial white mold-resistant dry bean cultivars exist.  -->
<!-- Field trials in some regions of the USA have shown the G122 cultivar is partially resistant while the Beryl cultivar is susceptible [@otto-hanson2011variation; @steadman2004identification]. -->
<!-- This puny paragraph needs a lot of work to more fully describe the operation of the wmn -->

<!-- Microsatellite markers (simple sequence repeats or SSR) are codominant, polymorphic and also able to detect multiple alleles per locus [@atallah2004high; @sirjusingh2001characterisation].
Reproducibility of microsatellites is higher compared other fingerprinting techniques such as random amplified polymorphic DNA (RAPD) [@atallah2004high].
High mutagenic rate and multiallelic nature of microsatellites render suitability for inferring population structure of organisms and their potential for outcrossing.
In the case of plant pathogens like *S. sclerotiorum* that is predominantly clonal and occasionally outcrossing, microsatellites have been used successfully to ascertain their ability to reproduce by sexual outcrossing [@atallah2004high; @sexton2004microsatellite] and also to detect differences among clonally propagated individuals.  -->

>>>>

Currently, there are no known resistant cultivars of dry bean commercially available [@otto-hanson2011variation].
Organized breeding efforts have been using a common-garden approach with white mold screening nurseries in dry bean production areas across the United States with additional sites in Mexico, France, and Australia [@steadman2003identification; @steadman2004identification; @steadman2005identification; @steadman2006identification; @otto2007identification]. 
These white mold screening nurseries use no chemical or cultural treatments against *S. sclerotiorum* and employ standardized protocols for screening new cultivars for resistance to white mold. 
It has previously been shown that aggressiveness (the severity of disease symptoms on the host) is significantly different across white mold screening nursery sites in separate geographic regions [@otto-hanson2011variation].
The genetic structure and mode of reproduction in these populations, however, is currently unknown.

Understanding genetic relationships and reproduction behavior of *S. sclerotiorum* populations is beneficial for breeders seeking to develop new resistant cultivars for worldwide deployment [@mcdonald2002pathogen; @milgroom1996recombination].
In particular, genetically diverse populations with high rates of sexual reproduction are more likely to overcome host resistance. 
Most populations of *S. sclerotiorum* are predominantly clonal with low genetic diversity and have a large degree of population fragmentation [@attanayake2012genetic; @carbone2001multilocus; @cubeta1997clonality; @ekins2011population; @kohli1998random; @kohli1995clonal]. 
Some studies, however have found populations that show signatures of sexual reproduction [@atallah2004high; @attanayake2013sclerotinia; @sexton2004microsatellite; @aldrich-wolfe2015genetic].

@@
